---
output: pdf_document
---
---
title: "Math HW6"
author: "Matthew Vanaman"
date: '`r format(Sys.Date(), "%m-%d-%Y")`'
monofont: Courier New
output:
  pdf_document: 
    latex_engine: xelatex
  word_document: default
header-includes: \usepackage{tikz,xcolor,listings,setspace}
indent: false
---
`r knitr::opts_knit$set(root.dir='..')`

```{r, include=FALSE}
library(ProjectTemplate); load.project()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
options(digits = 3, scipen = 999)
knitr::opts_chunk$set(engine.path = "/anaconda3/bin/python")
```

*All work and code are shown in the appendix.*

# 6.1

## (a)

$$f(x) = \int_{-1.96}^{1.96} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5^2} dx$$

### Left
```{r, echo=FALSE, fig.height=3}
# Variables
x <- seq(-3, 3, by=0.1)
norm <- dnorm(x, mean=0, sd=1)
df <- as.data.frame(cbind(x,norm))
# plot
ggplot(df, aes(x)) +
  geom_line(aes(y = norm)) +
  ylab("f(x)") + xlab("x") +
  labs(color = "Function") +
  theme_minimal() +
  scale_x_continuous(breaks=c(-1.96,1.96)) 
```

### Right
```{r, echo=FALSE, fig.height=3}
# Variables
x <- seq(-3, 3, by=0.1)
norm <- dnorm(x, mean=0, sd=1)
df <- as.data.frame(cbind(x,norm))
# plot
ggplot(df, aes(x)) +
  geom_line(aes(y = norm)) +
  ylab("f(x)") + xlab("x") +
  labs(color = "Function") +
  theme_minimal() +
  scale_x_continuous(breaks=c(-1.96,1.96)) 
```

### Midpoint
```{r, echo=FALSE, fig.height=3}
# Variables
x <- seq(-3, 3, by=0.1)
norm <- dnorm(x, mean=0, sd=1)
df <- as.data.frame(cbind(x,norm))
# plot
ggplot(df, aes(x)) +
  geom_line(aes(y = norm)) +
  ylab("f(x)") + xlab("x") +
  labs(color = "Function") +
  theme_minimal() +
  scale_x_continuous(breaks=c(-1.96,1.96)) 
```

### Trapezoidal
```{r, echo=FALSE, fig.height=3}
# Variables
x <- seq(-3, 3, by=0.1)
norm <- dnorm(x, mean=0, sd=1)
df <- as.data.frame(cbind(x,norm))
# plot
ggplot(df, aes(x)) +
  geom_line(aes(y = norm)) +
  ylab("f(x)") + xlab("x") +
  labs(color = "Function") +
  theme_minimal() +
  scale_x_continuous(breaks=c(-1.96,1.96)) 
```


## (b)

## Left

**Answer:** \newline
$A \approx \dfrac{0.392}{\sqrt{2 \pi}}  \Bigg( \sum_{i=0}^{9} e^{-0.5^2 \Big(0.392i - 1.96\Big)^2}\Bigg).$\newline


**Work:** \newline
The left Riemann Sum formula is:
$$A \approx \sum_{i=0}^{n-1} f\Big(a + \dfrac {(b-a)i}{n} \Big).$$
We want to find the sums for 10 partitions between -1.96 and 1.96. Which means that n = 10, a = -1.96, and  = 1.96. We also plug in the pdf for f. The Reimann Sum formula gets substituted in for x in f(x):
$$A \approx  \sum_{i=0}^{n-1} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5\Bigg(a + \frac {(b-a)i}{n} \Big( \frac{b - a}{n}\Big)\Bigg)^2} ,$$
$$\approx  \sum_{i=0}^{10-1} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Bigg(-1.96 + \frac {(1.96-(-1.96))i}{10} \Big(  \frac{1.96 - (-1.96)}{10}\Big)\Bigg)^2} ,$$
$$\approx \sum_{i=0}^{9} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Big((0.392i - 1.96)  0.392 \Big)^2} ,$$

0.392 does not depend on index, and neither does the constant $\frac{1}{\sqrt(2\pi)}$:

$$\approx \dfrac{1}{\sqrt{2 \pi}} (0.392) \Bigg( \sum_{i=0}^{9}  e^{-0.5 \Big((0.392i - 1.96)  0.392 \Big)^2}\Bigg),$$
$$\approx \dfrac{0.392}{\sqrt{2 \pi}}  \Bigg( \sum_{i=0}^{9} e^{-0.5^2 \Big(0.392i - 1.96\Big)^2}\Bigg).$$

```{r}
# calculation:
(1.96 + 1.96) / 10
```

## Right:

**Answer:** \newline
$A \approx \dfrac{0.392}{\sqrt{2 \pi}}  \Bigg( \sum_{i=1}^{10} e^{-0.5^2 \Big(0.392i - 1.96\Big)^2}\Bigg).$\newline

**Work:** \newline
Do the same thing here except the index is from $1$ to $n$ instead of from $0$ to $n-1$.
$$A \approx  \Bigg( \sum_{i=1}^{10} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Bigg(-1.96 + \frac {1.96-(-1.96)}{10} i \Big(  \frac{1.96 - (-1.96)}{10}\Big)\Bigg)^2}  \Bigg),$$
$$\approx \sum_{i=1}^{10} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Big((0.392i - 1.96)  0.392 \Big)^2},$$
$$\approx \dfrac{0.392}{\sqrt{2 \pi}}  \Bigg( \sum_{i=1}^{10} e^{-0.5^2 \Big(0.392i - 1.96\Big)^2}\Bigg).$$


## Midpoint:

**Answer:** \newline
$A \approx 0.392 \Bigg( \sum_{i=1}^{10} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Big(-2.16 + 0.392i)\Big)^2}  \Bigg).$ \newline

**Work:** \newline
$$A \approx \sum_{i=0}^{n-1} f\Big(a + \dfrac {(b-a)(i - 0.5)}{n} \Big).$$
We want to find the sums for 10 partitions between -1.96 and 1.96. Which means that n = 10, a = -1.96, and  = 1.96. We also plug in the pdf for f. The Reimann Sum formula gets substituted in for x in f(x):
$$A \approx  \Bigg( \sum_{i=0}^{n-1} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5\Bigg(a + \frac {(b-a)(i - 0.5)}{n} \Big( \frac{b - a}{n}\Big)\Bigg)^2} \Bigg),$$
$$\approx  \Bigg( \sum_{i=1}^{10} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Bigg(-1.96 + \frac {1.96-(-1.96)}{10} \Big(\frac{(i - 0.5)}{1}\Big) \Big(  \frac{1.96 - (-1.96)}{10}\Big)\Bigg)^2}  \Bigg),$$
$$\approx  \Bigg( \sum_{i=1}^{10} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Big(-1.96 + 0.392 (i - 0.5) (0.392)\Big)^2}  \Bigg),$$
$$\approx 0.392 \Bigg( \sum_{i=1}^{10} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Big(-1.96 + 0.392i - 0.196)\Big)^2}  \Bigg),$$
$$\approx 0.392 \Bigg( \sum_{i=1}^{10} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Big(-2.16 + 0.392i)\Big)^2}  \Bigg),$$
$$\approx 0.392 \Bigg( \sum_{i=1}^{10} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Big(-2.16 + 0.392i)\Big)^2}  \Bigg).$$

```{r}
# Calculations
0.392 * 0.5
-1.96 - 0.196
```

## Trapezoidal

**Answer:** \newline
$A \approx  \Big(\dfrac {0.196}{\sqrt{2\pi}}\Big) \sum_{i=0}^{9} \dfrac{ e^{-0.5(0.392i - 1.96)^2} + e^{-0.5(0.392i - 0.564)^2} } {2}.$ \newline

**Work:**
$$A \approx \sum_{i=0}^{n-1} \dfrac {b - a}{n} \dfrac{\dfrac{1}{\sqrt{2 \pi}} e^{  \Big(a + \frac {(b-a)i}{n} \Big)^2} + \dfrac {1}{\sqrt{2\pi}} e^{  \Big(a + \frac {(b-a)i+1}{n} \Big)^2} } {2} \dfrac {b-a}{n},$$
$$\approx \sum_{i=0}^{10-1} \dfrac {1.96 - (-1.96)}{10} \dfrac{\dfrac{1}{\sqrt{2 \pi}} e^{  \Big((-1.96) + \frac {(1.96-(-1.96))i}{10} \Big)^2} + \dfrac {1}{\sqrt{2\pi}} e^{  \Big((-1.96) + \frac {(1.96-(-1.96))i+1}{10} \Big)^2} } {2},$$
$$\approx \sum_{i=0}^{9} (0.392) \dfrac{\dfrac{1}{\sqrt{2 \pi}} e^{-0.5(0.392i - 1.96)^2} + \dfrac {1}{\sqrt{2\pi}} e^{-0.5(0.392i - 0.564)^2} } {2}, $$
$$\approx \dfrac {0.392}{2} \Big(\dfrac {1}{\sqrt{2\pi}}\Big) \sum_{i=0}^{9} \dfrac{ e^{-0.5(0.392i - 1.96)^2} + e^{-0.5(0.392i - 0.564)^2} } {2}, $$
$$\approx  \Big(\dfrac {0.196}{\sqrt{2\pi}}\Big) \sum_{i=0}^{9} \dfrac{ e^{-0.5(0.392i - 1.96)^2} + e^{-0.5(0.392i - 0.564)^2} } {2}.$$

```{r}
0.392 /2
```


## (c)

### Left

**Answer:** \newline
0.947 \newline

**Work:** \newline

Just plug in to x each level of n from 0 - 9:
$$A \approx \dfrac{0.392}{\sqrt{2 \pi}}  \Bigg( \sum_{i=0}^{9} e^{-0.5^2 \Big(0.392i - 1.96\Big)^2}\Bigg).$$
```{r}
e <- exp(1) # get euler's number
(0.392 / (sqrt(2 * pi))) * (e^(-0.5 * ((0.392 * 0) - 1.96)^2)   + 
                              e^(-0.5 * ((0.392 * 1) - 1.96)^2) + 
                              e^(-0.5 * ((0.392 * 2) - 1.96)^2) + 
                              e^(-0.5 * ((0.392 * 3) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 4) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 5) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 6) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 7) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 8) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 9) - 1.96)^2)
                            )
```

### Right

**Answer:** \newline
0.947 \newline

** Work:** \newline

Same thing, except runs from 1-10 instead of 0-9. Not surprising that it gives you the same answer:
```{r}
e <- exp(1) # get euler's number
(0.392 / (sqrt(2 * pi))) * (e^(-0.5 * ((0.392 * 1) - 1.96)^2)   +
                              e^(-0.5 * ((0.392 * 2) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 3) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 4) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 5) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 6) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 7) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 8) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 9) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 10) - 1.96)^2)
                            )
```

### Midpoint

**Answer:** \newline
0.951 \newline

**Work:** \newline

$$A \approx 0.392 \Bigg( \sum_{i=1}^{10} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Big(-2.16 + 0.392i)\Big)^2}  \Bigg).$$

```{r}
e <- exp(1) # get euler's number
(0.392 / (sqrt(2 * pi))) * (e^(-0.5 * ((0.392 * 1) - 2.16)^2)   + 
                              e^(-0.5 * ((0.392 * 2) - 2.16)^2) + 
                              e^(-0.5 * ((0.392 * 3) - 2.16)^2) +
                              e^(-0.5 * ((0.392 * 4) - 2.16)^2) +
                              e^(-0.5 * ((0.392 * 5) - 2.16)^2) +
                              e^(-0.5 * ((0.392 * 6) - 2.16)^2) +
                              e^(-0.5 * ((0.392 * 7) - 2.16)^2) +
                              e^(-0.5 * ((0.392 * 8) - 2.16)^2) +
                              e^(-0.5 * ((0.392 * 9) - 2.16)^2) +
                              e^(-0.5 * ((0.392 * 10) - 2.16)^2)
                            )
```

### Trapezoidal

**Answer:** \newline
0.947 \newline

**Work:** \newline

$$A \approx  \Big(\dfrac {0.196}{\sqrt{2\pi}}\Big) \sum_{i=0}^{9} \dfrac{ e^{-0.5(0.392i - 1.96)^2} + e^{-0.5(0.392i - 0.564)^2} } {2}.$$

```{r}
e <- exp(1) # get euler's number
(0.196 / (sqrt(2 * pi))) * (e^(-0.5 * ((0.392 * 0) - 1.96)^2)       +
                              e^(-0.5 * (0.392 * (0 + 1) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 1) - 1.96)^2)     +
                              e^(-0.5 * (0.392 * (1 + 1) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 2) - 1.96)^2)     +
                              e^(-0.5 * (0.392 * (2 + 1) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 3) - 1.96)^2)     +
                              e^(-0.5 * (0.392 * (3 + 1) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 4) - 1.96)^2)     +
                              e^(-0.5 * (0.392 * (4 + 1) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 5) - 1.96)^2)     +
                              e^(-0.5 * (0.392 * (5 + 1) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 6) - 1.96)^2)     +
                              e^(-0.5 * (0.392 * (6 + 1) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 7) - 1.96)^2)     +
                              e^(-0.5 * (0.392 * (7 + 1) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 8) - 1.96)^2)     +
                              e^(-0.5 * (0.392 * (8 + 1) - 1.96)^2) + 
                              e^(-0.5 * ((0.392 * 9) - 1.96)^2)     +
                              e^(-0.5 * (0.392 * (9 + 1) - 1.96)^2)
                            )
```

# 6.2

## (a)

**Answer:** \newline 
\newline
$= \dfrac {x^{101}} {101} + 3 e^x - 7 \dfrac {4^x}{1.39} + c.$ \newline

**Work:**\newline

$$F(x) = \int x^{100} + 3e^x - 7(4^x)dx,$$
$$= \int x^{100}dx + \int 3e^xdx - \int 7(4^x)dx,$$
$$= \dfrac {x^{100 + 1}} {100 + 1} + \int 3e^xdx - \int 7(4^x)dx,$$
$$= \dfrac {x^{101}} {101} + 3 \int e^xdx - 7 \int (4^x)dx,$$
The constant will cancel out, so leave it out.
$$= \dfrac {x^{101}} {101} + 3 e^x - 7 \dfrac {4^x}{\ln(4)} + c,$$
$$= \dfrac {x^{101}} {101} + 3 e^x - 7 \dfrac {4^x}{1.39} + c.$$

## (b)

**Answer:**
```{r, echo=FALSE}
cat((9^(3/2) * 10/3 - 1/9^3) - (1^(3/2) * 10/3 - 1/1^3))
```

**Work:**\newline
$$F(x) = \int_{1}^9 5 \sqrt{x} \ dx + \dfrac {3}{x^4} dx,$$
$$= \int_{1}^9 5 \sqrt{x} \ dx + \int \dfrac {3}{x^4} dx,$$
$$= 5 \int_{1}^9 \sqrt{x} \ dx + 3 \int \dfrac {1}{x^4} dx,$$
$$= 5 \int_{1}^9 x^{1/2} \ dx + 3 \int x^{-4} dx,$$
$$= 5 \dfrac {x^{1/2 + 1}} {\frac {1} {2} + 1}   + 3 \dfrac {x^{-4 + 1}}{-4 + 1},$$
$$= 5 \dfrac {x^{3/2}} {\frac {3} {2}} + 3 \dfrac {x^{-3}}{-3},$$
$$= 5 \Big(x^{3/2}\dfrac {2}{3}\Big) + 3 \Big( \dfrac {-1}{3}x^{-3} \Big),$$
$$= (x^{3/2}) \dfrac {10}{3} + (3) \dfrac {1}{x^3}.$$
From here, plug in the bounds and take the difference:
```{r}
(9^(3/2) * 10/3 - 1/9^3) - (1^(3/2) * 10/3 - 1/1^3)
```

## (c)

**Answer:**
```{r, echo=FALSE}
cat(12 * (0 + 1/2))
```

**Work:**
$$F(x) = \int_2^\infty \dfrac {12} {x^2} dx,$$
$$= \lim_{k \to \infty} \int_2^k \dfrac {12} {x^2} dx,$$
$$= 12 \lim_{k \to \infty} \int_2^k x^{-2} dx,$$
$$= 12 \lim_{k \to \infty} \int_2^k \dfrac {x^{-2+1}}{-2  +1} dx,$$
$$= 12 \lim_{k \to \infty} \dfrac {-1}{x}.$$
Plug in the bounds (plugging for the $-1/x$ since it approaches zero the higher x gets):
```{r}
12 * (0 + 1/2)
```

## (d)

**Answer:**\newline
$2y^2.$ \newline


**Work:**
$$F(x) = \dfrac {d}{dy} \int_{-3}^{y^2} \sqrt{x} \ dx,$$
$$= \sqrt{y^2} \ \dfrac {d}{dy} y^2,$$
Power rule...
$$= \sqrt{y^2} \ 2y,$$
$$= 2y^2.$$

## (e)

**Answer:**\newline
$- e^{\sqrt{z} + \ln(z)} \Big(\dfrac{1}{2 \sqrt{z}} + \dfrac {1}{z}\Big).$ \newline


**Work:**

$$F(x) = \dfrac {d}{dz} \int_{\sqrt{z} + \ln(z)}^{10} e^x dx,$$
$$= -1 \times \dfrac {d}{dz} \int_{\sqrt{z} + \ln(z)}^{10} e^x dx,$$
$$= -\dfrac {d}{dz} \int_{10}^{\sqrt{z} + \ln(z)} e^x dx,$$
$$= - e^{\sqrt{z} + \ln(z)} \dfrac{d}{dy} \sqrt{z} + \ln(z),$$
$$= - e^{\sqrt{z} + \ln(z)} \Big(\dfrac{1}{2 \sqrt{z}} + \dfrac {1}{z}\Big).$$

# 6.3a

**Answer:**\newline \newline
$\dfrac{(5x^{10} - 25x^4 + 15x)^{8}}{40} + c.$ \newline

**Work:**\newline

$$\int (10x^9 - 20x^3 + 3)(5x^10 - 25x^4 + 15x)^7 dx,$$
$$\int (10x^9 - 20x^3 + 3)(5x^{10} - 25x^4 + 15x)^7 dx.$$
$$u = 5x^{10} - 25x^4 + 15x,$$
$$\dfrac {du}{dx} = \dfrac{d}{dx}(5x^{10} - 25x^4 + 15x)$$
$$\dfrac {du}{dx}  = \dfrac{d}{dx}(5x^{10}) - \dfrac{d}{dx}(25x^4) + \dfrac{d}{dx}(15x),$$
$$\dfrac {du}{dx} = 5(10x^9) - 25(4x^3) + 15,$$
$$\dfrac {du}{dx} = 50x^9 - 100x^3 + 15,$$
$$\dfrac {1}{5} \dfrac {du}{dx} = (50x^9 - 100x^3 + 15)\dfrac {1}{5},$$
$$\dfrac {1}{5}\dfrac {du}{dx} = 10x^{10} - 20x^4 + 5x,$$
$$\dfrac {1}{5}du = (10x^{10} - 20x^4 + 5x)dx.$$
Substitute in $u$ and $1/5\ du$:
$$\int (10x^9 - 20x^3 + 3)(5x^{10} - 25x^4 + 15x)^7 dx,$$
$$\int (u)^7(\dfrac{1}{5}du),$$
$$\dfrac{1}{5}\int (u)^7(du),$$
$$\dfrac{1}{5}\int \dfrac{u^{8}}{8} + c,$$
$$\dfrac{u^{8}}{40} + c,$$
$$\dfrac{(5x^{10} - 25x^4 + 15x)^{8}}{40} + c.$$

# 6.4a

**Answer:** 
```{r, echo=FALSE}
cat(((2/3 * 4 * ((4+5)^(3/2))) - 2/3 * (2/5 * (4 + 5)^(5/2))) -
  
  ((2/3 * 1 * ((1+5)^(3/2))) - 2/3 * (2/5 * (1 + 5)^(5/2))))
```


$$\int_1^4 x \sqrt{x + 5} \ dx,$$
We want to get to:
$$\int u \ dv = uv - \int v \ du.$$
Use some u-substution:
$$u = x, \ dv = \sqrt{x + 5} \ dx.$$
$$\dfrac{du}{dx} = 1,\ \text{therefore} \ du = dx.$$
$$w = x + 5,$$ 
$$\dfrac{dw}{dx}(x+5) = 1 + 0 = 1,$$
$$dw = dx.$$
$$\int \sqrt{w} \ dw,$$
$$\int w^{\frac{1}{2}} \ (1),$$
$$\dfrac{w^{{\frac{1}{2} + 1}}} {\frac{1}{2} + 1},$$
$$\dfrac{w^{{\frac{3}{2}}}} {\frac{3}{2}},$$
$$ {\frac{2}{3}} w^{{\frac{3}{2}}},$$
$$ {\frac{2}{3}} (x + 5)^{{\frac{3}{2}}} = v.$$
Plug back into:
$$\int u \ dv = uv - \int v \ du,$$
with \newline
$u = x.$ \newline
$dv = \sqrt{x + 5} \ dx.$ \newline
$du = dx = dw.$ \newline
$w = x + 5.$ \newline
$v = {\frac{2}{3}} (x + 5)^{{\frac{3}{2}}}.$

$$\int_1^4 x \ \sqrt{x + 5} \ dx = x \ \Big({\frac{2}{3}} (x + 5)^{{\frac{3}{2}}}\Big) - \int_1^4 {\frac{2}{3}} (x + 5)^{{\frac{3}{2}}} \ dx,$$
$$\int_1^4 x \ \sqrt{x + 5} \ dx = x \ \Big({\frac{2}{3}} (x + 5)^{{\frac{3}{2}}}\Big) -{\frac{2}{3}}  \int_1^4  (x + 5)^{{\frac{3}{2}}} \ dx.$$
Before we can start plugging in numbers, gotta take care of the remaining integral:
$$ \int (x + 5)^{{\frac{3}{2}}} \ dx,$$
$$ \int (x + 5)^{{\frac{3}{2}}} \ dx,$$
$$u = x + 5,$$
$$ \int (u)^{{\frac{3}{2}}} \ dx,$$
$$ \int \dfrac{(u)^{{\frac {3} {2} + 1}}}{\frac {3} {2} + 1} \ dx,$$
$$\dfrac{(u)^{{\frac {5} {2}}}}{\frac {5} {2}},$$
$$\frac {2} {5}(u)^{{\frac {5} {2}}},$$
$$\frac {2} {5}(x + 5)^{{\frac {5} {2}}}.$$
So in the end you have:
$$\int_1^4 x \ \sqrt{x + 5} \ dx = x \ \Big({\frac{2}{3}} (x + 5)^{{\frac{3}{2}}}\Big) -{\frac{2}{3}} \Big( \frac {2} {5}(x + 5)^{{\frac {5} {2}}}\Big).$$
Plug in the bounds:
```{r}
((2/3 * 4 * ((4+5)^(3/2))) - 2/3 * (2/5 * (4 + 5)^(5/2))) -
  
  ((2/3 * 1 * ((1+5)^(3/2))) - 2/3 * (2/5 * (1 + 5)^(5/2)))
```


# 6.5a

**Answer:** 
```{r, echo=FALSE}
e <- exp(1)
cat(167 / 3 - (log(8))^3 / 3 + e^(167^(1/3)) - e^(log(8)))
```

**Work:**
Second theorem:
$$\int_a^b f(x) \ dx = F(x) \Bigg\rvert_a^b = F(b) - F(a).$$
$$F(x) = \int_{\ln(8)}^{\sqrt[3]{167}} x^2 + e^x \ dx,$$
$$ = \int_{\ln(8)}^{ \sqrt[3]{167}} x^2 + \int_{\ln(8)}^{\sqrt[3]{167}}e^x \ dx,$$
$$ = \dfrac{x^{3}} {3} + e^x + c,$$
$$ = \Big(\dfrac{x^{3}} {3} - \dfrac{x^{3}} {3} \Big) + \Big( (e^x + c) - (e^x + c) \Big),$$
the constant cancels out
$$ = \Big(\dfrac{(\sqrt[3]{167})^{3}} {3} - \dfrac{\ln(8)^{3}} {3} \Big) + \Big( e^{\sqrt[3]{167}} - e^{\ln(8)} \Big),$$
$$ = \Big(\dfrac{(\sqrt[3]{167})^{3}} {3} - \dfrac{\ln(8)^{3}} {3} \Big) + \Big( e^{\sqrt[3]{167}} - e^{\ln(8)} \Big),$$
$$= 291. $$
```{r}
e <- exp(1)
167 / 3 - (log(8))^3 / 3 + e^(167^(1/3)) - e^(log(8))
```


# 6.6

## (a)

**Answer:** \newline
A function is a PDF if the function is never less than 0 and the total area under the curve (over the domain of the function) is equal to 1.\newline
Step 1: show that function is never less than 0:
```{r, results = "hold"}
(3 * sqrt(0) / 2)
(3 * sqrt(1) / 2)
(3 * sqrt(100) / 2)
```
Function is never less than 0! \newline
Step 2: show that the total area under the curve over the domain is equal to 1.\newline
To get the area under the curve, integrate the function (it'd be a definite integral):
$$\int_0^1 \dfrac{3 \sqrt{x}} {2} dx,$$
$$ \dfrac{3} {2}  \int_0^1 x^{1/2} \ dx,$$
$$ \dfrac{3} {2}  \int_0^1 \dfrac{x^{\frac{1}{2} + 1}} {\frac {1}{2} + 1} \ dx,$$
$$ \dfrac{3} {2}  \int_0^1 \dfrac{x^{3/2}} {\frac {3}{2}} \ dx,$$
$$ \dfrac{3} {2} \Big(\dfrac{2}{3} x^{3/2}\Big) \ dx,$$
$$x^{3/2}.$$

Plug in the values of the domain and get the area under the curve within the domain of the function:
```{r}
1^(3/2) - 0^(3/2)
```
Because this function satisfies both criteria, it is a PDF.

## (b)
**Answer:**
```{r, echo=FALSE}
cat("A:", (1^(3/2) - 0.9^(3/2)),"\n")
cat("B:", 0.9^(3/2) - 0.8^(3/2),"\n")
cat("C:", 0.8^(3/2) - 0.7^(3/2),"\n")
cat("D:", 0.7^(3/2) - 0.6^(3/2),"\n") 
cat("F:", 0.6^(3/2) - 0.0^(3/2),"\n")
```

**Work:** \newline
We got the integral, so now we just need to plug in the domains of all the grades. Since x runs from 0 to 1, we know that the range of e.g. an A is from 0.9 to 1.0 on the domain.
```{r, results = "hide"}
(1^(3/2) - 0.9^(3/2))
0.9^(3/2) - 0.8^(3/2)
0.8^(3/2) - 0.7^(3/2)
0.7^(3/2) - 0.6^(3/2)
0.6^(3/2) - 0.0^(3/2)
```

## (c)
**Answer:**
```{r, echo=FALSE}
cat(3/5 * 1^(5/2) - 3/5 * 0^(5/2))
```

**Work:** \newline
We have the formula for the first moment:
$$E(x) = \int_{-\infty}^\infty x f(x) \ dx.$$
We just want to plug in the information we have:\newline
- $f(x) = \frac {3 \sqrt{x}} {2}$ \newline
- the domain of x: [0,1].
$$E(x) = \int_0^1 x \frac {3 \sqrt{x}} {2} \ dx,$$
$$= \int_0^1 \frac {3x (x^{1/2})} {2} \ dx,$$
$$= \int_0^1 \frac{3}{2} x^{3/2} \ dx,$$
$$= \frac{3}{2} \int_0^1 \dfrac{x^{3/2 + 1}} {\frac {3}{2} + 1} \ dx,$$
$$= \frac{3}{2} \Big( {\frac {2}{5}} x^{5/2} \Big),$$
$$= {\frac {3}{5}} x^{5/2}$$

Plug in the domain to the first moment formula:
```{r}
3/5 * 1^(5/2) - 3/5 * 0^(5/2)
```

## (d)

**Answer:**
```{r, echo = FALSE}
cat("Variance:", (0.429 - 0.6^2) * 100, "\n",
"Standard deviation:", sqrt(0.429 - 0.6^2) * 100)
```

**Work:** \newline
Second moment:
$$V(x) = E(x^2) - E(x)^2.$$
- $E(x)^2 = 0.6^2$\newline
- $E(x^2) = \ ?$
$$\int_0^1 x^2 \dfrac {3 \sqrt{x}}{2} dx,$$
$$\dfrac {3}{2} \int_0^1 x^2 \sqrt{x} \ dx,$$
$$\dfrac {3}{2} \int_0^1 x^2 (x^{1/2}) \ dx,$$
$$\dfrac {3}{2} \int_0^1 x^{5/2} \ dx,$$
$$\dfrac {3}{2} \int_0^1 \dfrac{x^{7/2}}{\frac{7}{2}} \ dx,$$
$$\dfrac {3}{2} \frac{7}{2} x^{7/2},$$
$$\dfrac {3}{7} x^{7/2},$$
$$\dfrac {3}{7} (1^{7/2}) - \dfrac {3}{7} (0^{7/2}),$$
$$= 0.429.$$
```{r}
# calculation
3/7 * 1^(7/2) - 3/7 * 0^(7/2)
```
Multiply by 100 to get onto 0-100 scale \newline
Variance:
$V(x) = E(x^2) - E(x)^2:$
```{r}
(0.429 - 0.6^2) * 100
```
Standard deviation:
```{r}
sqrt(0.429 - 0.6^2) * 100
```
