---
title: "Math HW6"
author: "Matthew Vanaman"
date: '`r format(Sys.Date(), "%m-%d-%Y")`'
monofont: Courier New
output:
  pdf_document: 
    latex_engine: xelatex
  word_document: default
header-includes: \usepackage{tikz,xcolor,listings,setspace}
indent: false
---
`r knitr::opts_knit$set(root.dir='..')`

```{r, include=FALSE}
library(ProjectTemplate); load.project()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
options(digits = 3, scipen = 999)
knitr::opts_chunk$set(engine.path = "/anaconda3/bin/python")
```

6.2, 6.3a, 6.4a, 6.5a, 6.6

# 6.1

## (a)

$$f(x) = \int_{-1.96}^{1.96} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5^2} dx$$

### Left
```{r, echo=FALSE, fig.height=3}
# Variables
x <- seq(-3, 3, by=0.1)
norm <- dnorm(x, mean=0, sd=1)
df <- as.data.frame(cbind(x,norm))
# plot
ggplot(df, aes(x)) +
  geom_line(aes(y = norm)) +
  ylab("f(x)") + xlab("x") +
  labs(color = "Function") +
  theme_minimal() +
  scale_x_continuous(breaks=c(-1.96,1.96)) 
```

### Right
```{r, echo=FALSE, fig.height=3}
# Variables
x <- seq(-3, 3, by=0.1)
norm <- dnorm(x, mean=0, sd=1)
df <- as.data.frame(cbind(x,norm))
# plot
ggplot(df, aes(x)) +
  geom_line(aes(y = norm)) +
  ylab("f(x)") + xlab("x") +
  labs(color = "Function") +
  theme_minimal() +
  scale_x_continuous(breaks=c(-1.96,1.96)) 
```

### Midpoint
```{r, echo=FALSE, fig.height=3}
# Variables
x <- seq(-3, 3, by=0.1)
norm <- dnorm(x, mean=0, sd=1)
df <- as.data.frame(cbind(x,norm))
# plot
ggplot(df, aes(x)) +
  geom_line(aes(y = norm)) +
  ylab("f(x)") + xlab("x") +
  labs(color = "Function") +
  theme_minimal() +
  scale_x_continuous(breaks=c(-1.96,1.96)) 
```

### Trapezoidal
```{r, echo=FALSE, fig.height=3}
# Variables
x <- seq(-3, 3, by=0.1)
norm <- dnorm(x, mean=0, sd=1)
df <- as.data.frame(cbind(x,norm))
# plot
ggplot(df, aes(x)) +
  geom_line(aes(y = norm)) +
  ylab("f(x)") + xlab("x") +
  labs(color = "Function") +
  theme_minimal() +
  scale_x_continuous(breaks=c(-1.96,1.96)) 
```


## (b)

## Left

**Answer:** \newline
$A \approx \dfrac{0.392}{\sqrt{2 \pi}}  \Bigg( \sum_{i=0}^{9} e^{-0.5^2 \Big(0.392i - 1.96\Big)^2}\Bigg).$\newline


**Work:** \newline
The left Riemann Sum formula is:
$$A \approx \sum_{i=0}^{n-1} f\Big(a + \dfrac {(b-a)i}{n} \Big).$$
We want to find the sums for 10 partitions between -1.96 and 1.96. Which means that n = 10, a = -1.96, and  = 1.96. We also plug in the pdf for f. The Reimann Sum formula gets substituted in for x in f(x):
$$A \approx  \sum_{i=0}^{n-1} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5\Bigg(a + \frac {(b-a)i}{n} \Big( \frac{b - a}{n}\Big)\Bigg)^2} ,$$
$$\approx  \sum_{i=0}^{10-1} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Bigg(-1.96 + \frac {(1.96-(-1.96))i}{10} \Big(  \frac{1.96 - (-1.96)}{10}\Big)\Bigg)^2} ,$$
$$\approx \sum_{i=0}^{9} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Big((0.392i - 1.96)  0.392 \Big)^2} ,$$

0.392 does not depend on index, and neither does the constant $\frac{1}{\sqrt(2\pi)}$:

$$\approx \dfrac{1}{\sqrt{2 \pi}} (0.392) \Bigg( \sum_{i=0}^{9}  e^{-0.5 \Big((0.392i - 1.96)  0.392 \Big)^2}\Bigg),$$
$$\approx \dfrac{0.392}{\sqrt{2 \pi}}  \Bigg( \sum_{i=0}^{9} e^{-0.5^2 \Big(0.392i - 1.96\Big)^2}\Bigg).$$

```{r}
# calculation:
(1.96 + 1.96) / 10
```

## Right:

**Answer:** \newline
$A \approx \dfrac{0.392}{\sqrt{2 \pi}}  \Bigg( \sum_{i=1}^{10} e^{-0.5^2 \Big(0.392i - 1.96\Big)^2}\Bigg).$\newline

**Work:** \newline
Do the same thing here except the index is from $1$ to $n$ instead of from $0$ to $n-1$.
$$A \approx  \Bigg( \sum_{i=1}^{10} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Bigg(-1.96 + \frac {1.96-(-1.96)}{10} i \Big(  \frac{1.96 - (-1.96)}{10}\Big)\Bigg)^2}  \Bigg),$$
$$\approx \sum_{i=1}^{10} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Big((0.392i - 1.96)  0.392 \Big)^2},$$
$$\approx \dfrac{0.392}{\sqrt{2 \pi}}  \Bigg( \sum_{i=1}^{10} e^{-0.5^2 \Big(0.392i - 1.96\Big)^2}\Bigg).$$


## Midpoint:

**Answer:** \newline
$A \approx 0.392 \Bigg( \sum_{i=1}^{10} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Big(-2.16 + 0.392i)\Big)^2}  \Bigg).$ \newline

**Work:** \newline
$$A \approx \sum_{i=0}^{n-1} f\Big(a + \dfrac {(b-a)(i - 0.5)}{n} \Big).$$
We want to find the sums for 10 partitions between -1.96 and 1.96. Which means that n = 10, a = -1.96, and  = 1.96. We also plug in the pdf for f. The Reimann Sum formula gets substituted in for x in f(x):
$$A \approx  \Bigg( \sum_{i=0}^{n-1} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5\Bigg(a + \frac {(b-a)(i - 0.5)}{n} \Big( \frac{b - a}{n}\Big)\Bigg)^2} \Bigg),$$
$$\approx  \Bigg( \sum_{i=1}^{10} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Bigg(-1.96 + \frac {1.96-(-1.96)}{10} \Big(\frac{(i - 0.5)}{1}\Big) \Big(  \frac{1.96 - (-1.96)}{10}\Big)\Bigg)^2}  \Bigg),$$
$$\approx  \Bigg( \sum_{i=1}^{10} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Big(-1.96 + 0.392 (i - 0.5) (0.392)\Big)^2}  \Bigg),$$
$$\approx 0.392 \Bigg( \sum_{i=1}^{10} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Big(-1.96 + 0.392i - 0.196)\Big)^2}  \Bigg),$$
$$\approx 0.392 \Bigg( \sum_{i=1}^{10} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Big(-2.16 + 0.392i)\Big)^2}  \Bigg),$$
$$\approx 0.392 \Bigg( \sum_{i=1}^{10} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Big(-2.16 + 0.392i)\Big)^2}  \Bigg).$$

```{r}
# Calculations
0.392 * 0.5
-1.96 - 0.196
```

## Trapezoidal

**Answer:** \newline
$A \approx  \Big(\dfrac {0.196}{\sqrt{2\pi}}\Big) \sum_{i=0}^{9} \dfrac{ e^{-0.5(0.392i - 1.96)^2} + e^{-0.5(0.392i - 0.564)^2} } {2}.$ \newline

**Work:**
$$A \approx \sum_{i=0}^{n-1} \dfrac {b - a}{n} \dfrac{\dfrac{1}{\sqrt{2 \pi}} e^{  \Big(a + \frac {(b-a)i}{n} \Big)^2} + \dfrac {1}{\sqrt{2\pi}} e^{  \Big(a + \frac {(b-a)i+1}{n} \Big)^2} } {2} \dfrac {b-a}{n},$$
$$\approx \sum_{i=0}^{10-1} \dfrac {1.96 - (-1.96)}{10} \dfrac{\dfrac{1}{\sqrt{2 \pi}} e^{  \Big((-1.96) + \frac {(1.96-(-1.96))i}{10} \Big)^2} + \dfrac {1}{\sqrt{2\pi}} e^{  \Big((-1.96) + \frac {(1.96-(-1.96))i+1}{10} \Big)^2} } {2},$$
$$\approx \sum_{i=0}^{9} (0.392) \dfrac{\dfrac{1}{\sqrt{2 \pi}} e^{-0.5(0.392i - 1.96)^2} + \dfrac {1}{\sqrt{2\pi}} e^{-0.5(0.392i - 0.564)^2} } {2}, $$
$$\approx \dfrac {0.392}{2} \Big(\dfrac {1}{\sqrt{2\pi}}\Big) \sum_{i=0}^{9} \dfrac{ e^{-0.5(0.392i - 1.96)^2} + e^{-0.5(0.392i - 0.564)^2} } {2}, $$
$$\approx  \Big(\dfrac {0.196}{\sqrt{2\pi}}\Big) \sum_{i=0}^{9} \dfrac{ e^{-0.5(0.392i - 1.96)^2} + e^{-0.5(0.392i - 0.564)^2} } {2}.$$

```{r}
0.392 /2
```


## (c)

### Left

**Answer:** \newline
0.947 \newline

**Work:** \newline

Just plug in to x each level of n from 0 - 9:
$$A \approx \dfrac{0.392}{\sqrt{2 \pi}}  \Bigg( \sum_{i=0}^{9} e^{-0.5^2 \Big(0.392i - 1.96\Big)^2}\Bigg).$$
```{r}
e <- exp(1) # get euler's number
(0.392 / (sqrt(2 * pi))) * (e^(-0.5 * ((0.392 * 0) - 1.96)^2)   + 
                              e^(-0.5 * ((0.392 * 1) - 1.96)^2) + 
                              e^(-0.5 * ((0.392 * 2) - 1.96)^2) + 
                              e^(-0.5 * ((0.392 * 3) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 4) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 5) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 6) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 7) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 8) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 9) - 1.96)^2)
                            )
```

### Right

**Answer:** \newline
0.947 \newline

** Work:** \newline

Same thing, except runs from 1-10 instead of 0-9. Not surprising that it gives you the same answer:
```{r}
e <- exp(1) # get euler's number
(0.392 / (sqrt(2 * pi))) * (e^(-0.5 * ((0.392 * 1) - 1.96)^2)   +
                              e^(-0.5 * ((0.392 * 2) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 3) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 4) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 5) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 6) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 7) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 8) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 9) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 10) - 1.96)^2)
                            )
```

### Midpoint

**Answer:** \newline
0.951 \newline

**Work:** \newline

$$A \approx 0.392 \Bigg( \sum_{i=1}^{10} \dfrac{1}{\sqrt{2 \pi}} e^{-0.5 \Big(-2.16 + 0.392i)\Big)^2}  \Bigg).$$

```{r}
e <- exp(1) # get euler's number
(0.392 / (sqrt(2 * pi))) * (e^(-0.5 * ((0.392 * 1) - 2.16)^2)   + 
                              e^(-0.5 * ((0.392 * 2) - 2.16)^2) + 
                              e^(-0.5 * ((0.392 * 3) - 2.16)^2) +
                              e^(-0.5 * ((0.392 * 4) - 2.16)^2) +
                              e^(-0.5 * ((0.392 * 5) - 2.16)^2) +
                              e^(-0.5 * ((0.392 * 6) - 2.16)^2) +
                              e^(-0.5 * ((0.392 * 7) - 2.16)^2) +
                              e^(-0.5 * ((0.392 * 8) - 2.16)^2) +
                              e^(-0.5 * ((0.392 * 9) - 2.16)^2) +
                              e^(-0.5 * ((0.392 * 10) - 2.16)^2)
                            )
```

### Trapezoidal

**Answer:** \newline
0.947 \newline

**Work:** \newline

$$A \approx  \Big(\dfrac {0.196}{\sqrt{2\pi}}\Big) \sum_{i=0}^{9} \dfrac{ e^{-0.5(0.392i - 1.96)^2} + e^{-0.5(0.392i - 0.564)^2} } {2}.$$

```{r}
e <- exp(1) # get euler's number
(0.196 / (sqrt(2 * pi))) * (e^(-0.5 * ((0.392 * 0) - 1.96)^2)       +
                              e^(-0.5 * (0.392 * (0 + 1) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 1) - 1.96)^2)     +
                              e^(-0.5 * (0.392 * (1 + 1) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 2) - 1.96)^2)     +
                              e^(-0.5 * (0.392 * (2 + 1) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 3) - 1.96)^2)     +
                              e^(-0.5 * (0.392 * (3 + 1) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 4) - 1.96)^2)     +
                              e^(-0.5 * (0.392 * (4 + 1) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 5) - 1.96)^2)     +
                              e^(-0.5 * (0.392 * (5 + 1) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 6) - 1.96)^2)     +
                              e^(-0.5 * (0.392 * (6 + 1) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 7) - 1.96)^2)     +
                              e^(-0.5 * (0.392 * (7 + 1) - 1.96)^2) +
                              e^(-0.5 * ((0.392 * 8) - 1.96)^2)     +
                              e^(-0.5 * (0.392 * (8 + 1) - 1.96)^2) + 
                              e^(-0.5 * ((0.392 * 9) - 1.96)^2)     +
                              e^(-0.5 * (0.392 * (9 + 1) - 1.96)^2)
                            )
```

# 6.2

## (a)

**Answer:** \newline 
\newline
$= \dfrac {x^{101}} {101} + 3 e^x - 7 \dfrac {4^x}{1.39} + c.$ \newline

**Work:**\newline

$$F(x) = \int x^{100} + 3e^x - 7(4^x)dx,$$
$$= \int x^{100}dx + \int 3e^xdx - \int 7(4^x)dx,$$
$$= \dfrac {x^{100 + 1}} {100 + 1} + \int 3e^xdx - \int 7(4^x)dx,$$
$$= \dfrac {x^{101}} {101} + 3 \int e^xdx - 7 \int (4^x)dx,$$
$$= \dfrac {x^{101}} {101} + 3 e^x - 7 \dfrac {4^x}{\ln(4)} + c,$$
$$= \dfrac {x^{101}} {101} + 3 e^x - 7 \dfrac {4^x}{1.39} + c.$$
Should $3 * \int e^xdx$ become $3 * e^x + c$?

## (b)

**Answer:**
```{r, echo=FALSE}
cat((9^(3/2) * 10/3 - 1/9^3) - (1^(3/2) * 10/3 - 1/1^3))
```

**Work:**\newline
$$F(x) = \int_{1}^9 5 \sqrt{x} \ dx + \dfrac {3}{x^4} dx,$$
$$= \int_{1}^9 5 \sqrt{x} \ dx + \int \dfrac {3}{x^4} dx,$$
$$= 5 \int_{1}^9 \sqrt{x} \ dx + 3 \int \dfrac {1}{x^4} dx,$$
$$= 5 \int_{1}^9 x^{1/2} \ dx + 3 \int x^{-4} dx,$$
$$= 5 \dfrac {x^{1/2 + 1}} {\frac {1} {2} + 1}   + 3 \dfrac {x^{-4 + 1}}{-4 + 1},$$
$$= 5 \dfrac {x^{3/2}} {\frac {3} {2}} + 3 \dfrac {x^{-3}}{-3},$$
$$= 5 \Big(x^{3/2}\dfrac {2}{3}\Big) + 3 \Big( \dfrac {-1}{3}x^{-3} \Big),$$
$$= (x^{3/2}) \dfrac {10}{3} + (3) \dfrac {1}{x^3}.$$
From here, plug in the bounds and take the difference:
```{r}
(9^(3/2) * 10/3 - 1/9^3) - (1^(3/2) * 10/3 - 1/1^3)
```

