---
title: "Math HW8"
author: "Matthew Vanaman"
date: '`r format(Sys.Date(), "%m-%d-%Y")`'
monofont: Courier New
output:
  pdf_document: 
    latex_engine: xelatex
  word_document: default
header-includes: \usepackage{listings}
---
`r knitr::opts_knit$set(root.dir='..')`

```{r, include=FALSE, cache=TRUE}
library(ProjectTemplate); load.project()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
options(scipen = 999)
knitr::opts_chunk$set(engine.path = "/anaconda3/bin/python")
```

8.4, 8.8, 8.9, 8.10a, 8.11 (you should use a matrix language to do this problem, but be sure to show suitable intermediate steps)

## 8.3

I made the matrices here for checking work (see code in the appendix), but just to show they work as intended using A as an example:
```{r, echo=FALSE}
A <- matrix(c(-1,3,4), nrow=3, ncol=1)
B <- matrix(c(3,-2,1), nrow=3, ncol=1)
C <- matrix(c(3,-8,2,0,-4,6), nrow=2, ncol=3)
D <- matrix(c(6,-1,-3,-2,3,8), nrow=3, ncol=2)
# this function gets rid of the annoying row and  column names of the matrices
print.matrix <- function(m){
write.table(format(m, justify="right"),
            row.names=F, col.names=F, quote=F)
}
# This function makes r matrix output look pretty and consistent with latex
write_matex <- function(x) {
  begin <- "$$= \\begin{bmatrix}"
  end <- "\\end{bmatrix}.$$"
  X <-
    apply(x, 1, function(x) {
      paste(
        paste(x, collapse = "&"),
        "\\\\"
      )
    })
  writeLines(c(begin, X, end))
}
```
```{r}
print(A)
print(A * 3)
```


**\large (a):** 
$$3A - 2B,$$
```{r, echo=FALSE, results="asis"}
write_matex(3*A - 2*B)
```

**Work:**
If the number of rows and columns are the same, you can use scalar multiplication. 
$$3A = \begin{bmatrix} 3(-1) \\ 3(3) \\ 3(4) \end{bmatrix},$$

```{r, results="asis"}
write_matex(3*A)
```

$$2B = \begin{bmatrix} 2(3) \\ 2(-2) \\ 2(1) \end{bmatrix},$$
```{r, results="asis"}
write_matex(2*B)
```
$$3A - 2B,$$
```{r results="asis"}
write_matex(3*A - 2*B)
```

**\large (b):** $A \cdot B$ = -5. \newline

**Work:** \newline
Element-wise multiplication:\newline
$$A \times B,$$
$$= \begin{bmatrix}-1 \\ 3 \\ 4  \end{bmatrix} \times \begin{bmatrix} 3 \\ -2 \\ 1  \end{bmatrix},$$
$$= -1(3) + 3(-2) + 4(1),$$
$$-5.$$

**\large (c):**
$$A \times B,$$
```{r, echo=FALSE, results="asis"}
write_matex(A%o%B)
```


**Work:** \newline
Out-product rule: 
$$A \times B,$$
$$= \begin{bmatrix} -1(3) & -1(-2) & -1(1) \\ 3(3) & 3(-2) & 3(1) \\ 4(3) & 4(-2) & 4(1) \end{bmatrix},$$
```{r, results="asis"}
write_matex(A%o%B)
```

**\large (d):** 
$$CA,$$
```{r, echo=FALSE, results="asis"}
write_matex(C%*%A)
```

**Work:** \newline
Matrix multiplication, which is possible because the number of C's columns are equal to the number of A's rows: $C_{23} \ A_{32}$.
$$CA,$$
$$= \begin{bmatrix} 3(-1) + 2(3) + -4(4) \\ -8(-1) + 0(3) + 6(4) \end{bmatrix},$$
```{r, results="asis"}
write_matex(C%*%A)
```

**\large (e):** Because the dimensions don't work out ($B_{31} \ D_{32}$), we can't multiply.\newline

**\large (f):** 
$$B \otimes D,$$
```{r, echo=FALSE, results="asis"}
write_matex(kronecker(B,D, FUN="*"))
```


**Work:**
$$B \otimes D,$$
$$= \begin{bmatrix} 3 \ \ \small \begin{bmatrix} 6 & -2 \\ -1 & 3 \\ -3 & 8 \end{bmatrix} \\ -1 \small \begin{bmatrix}6 & -2 \\ -1 & 3 \\ -3 & 8 \end{bmatrix} \\ 1 \ \ \small \begin{bmatrix} 6 & -2 \\ -1 & 3 \\ -3 & 8 \end{bmatrix}   \end{bmatrix}$$
```{r, results="asis"}
write_matex(kronecker(B,D, FUN="*"))
```

**\large (g):** 
$$CD,$$
```{r, echo=FALSE, results="asis"}
write_matex(C%*%D)
```


**Work:** \newline
Same situation as (d), because the dimensions work out: $C_{23}$ and $D_{32}$.
$$CD,$$
$$= \begin{bmatrix} 3(6) + 2(1) + (-4)(-3) & 3(-2) + 2(3) + (-1)(8) \\ -8(6) + 0(-1) + 6(-3) & -8(-2) + 0(3) + 6(8) \end{bmatrix},$$
```{r, results="asis"}
write_matex(C%*%D)
```

**\large (h):** 
$$DC,$$
```{r, echo=FALSE, results="asis"}
write_matex(D%*%C)
```


**Work:** \newline
Yet again the same situation as (d), because the dimensions work out: $D_{32}$ and $C_{23}$.
$$DC,$$
$$= \begin{bmatrix} 6(3) + (-2)(-8) &  6(2) + (-2)0 & 6(-4) + (-2)6 \\ -1(3) + 3(-8) & -1(2) + 3(0) & -1(-4) + 3(6) \\ -3(3) + 8(-8) & -3(2) + 8(0) & -3(-4) + 8(6) \end{bmatrix},$$
```{r, results="asis"}
write_matex(C%*%D)
```

**\large (i):** 
$$C'C,$$
```{r, echo=FALSE, results="asis"}
write_matex(t(C)%*%C)
```

**Work:** \newline
First, transpose C:
```{r, results="asis"}
write_matex(t(C))
```
$$C'C,$$
$$= \begin{bmatrix} 3(3) + (-8)(-8) &  3(2) + (-8)0 & 3(-4) + (-8)6 \\ 2(3) + 0(-8) & 2(2) + 0(0) & 2(-4) + 0(6) \\ -4(3) + 6(-8) & -4(2) + 6(0) & -4(-4) + 6(6) \end{bmatrix},$$
```{r, results="asis"}
write_matex(t(C)%*%C)
```

