---
title: "Math HW7"
author: "Matthew Vanaman"
date: '`r format(Sys.Date(), "%m-%d-%Y")`'
monofont: Courier New
output:
  pdf_document: 
    latex_engine: xelatex
  word_document: default
header-includes: \usepackage{listings}
---
`r knitr::opts_knit$set(root.dir='..')`

```{r, include=FALSE, cache=TRUE}
library(ProjectTemplate); load.project()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
options(scipen = 999)
knitr::opts_chunk$set(engine.path = "/anaconda3/bin/python")
knitr::knit_hooks$set(document = function(x) {sub('\\usepackage[]{color}','\\usepackage{xcolor}', x, fixed = TRUE)})
np <- reticulate::import("numpy")
sympy <- reticulate::import_from_path("sympy", path = "/anaconda3/lib/python3.7/site-packages")
```

*All work and code are shown in the Appendix.* \newline

**\large 7.1a:** $f(a,b,c,d,e,f) = a \times b \times c \times d \times e \times f$ \newline

**\large 7.3a:** 5.

## 7.6 

**\large (a):** $\nabla f(xy) = \begin{bmatrix} y - 2x + 2 \\ x - 2y + 1 \end{bmatrix}.$ \newline

**\large (b):** Critical point = $(\dfrac{5}{3},\dfrac{4}{3}).$ \newline

**\large (c):** $\begin{bmatrix} -2 & 1 \\ 1 & -2 \end{bmatrix}.$ \newline

**\large (d):** At $(\dfrac {5} {3}, \dfrac {4}{3})$, the matrix is negative-definite, and at a local maximum. \newline

**\large 7.7:** The pair (111.11, 133.33) maximizes the function.

## 7.10

**\large (a):** 8.33.

**\large (b):** 27. 

**\large (c):** 15.

**\large (d):** -6. 

**\large (e):** -0.298. 
\newpage

\begin{center}
\newpage
\LARGE{Appendix}
\end{center}

## 7.1a
Work shown in answer.

## 7.3a
**Work:** \newline
Plug in 3 and 1:
```{r}
(3^2 - 3 * 1 + (6 * 1)^2) / (3 - 3 * 1)
```
So that doesn't work. Need to find some way to cancel out the denominator.

$$x^2 - xy + 6y^2,$$
Set y equal to 1:
$$= x^2 - (x \cdot 1) + (6 \cdot 1)^2,$$
$$= x^2 - x + 6,$$
$$= (x - 3)(x + 2).$$
Plug y back in:
$$= (x - 3y)(x + 2y).$$
Check work:
```{python}
from __future__ import division
from sympy import *
x, y = symbols('x y')
print(expand((x - 3*y) * (x + 2*y)))
```
plug back into original expression:
$$\dfrac{(x - 3y)(x + 2y)}{x - 3y},$$
$$= x + 2y.$$
Plugging in x and y values:
```{r}
3 + 2 * 1
```

## 7.6

### (a)
**Work:**\newline
Gradient = first partial derivative. Need to find one for x and one for y, and construct a matrix where row 1 is the first partial derivative with respect to x and row 2 is the first partial derivative with respect to y.\newline
With respect to x:
$$\dfrac{\partial}{\partial x} (-x^2 + xy - y^2 + 2x  + y),$$
$$= \dfrac{\partial}{\partial x} (-x^2) + \dfrac{\partial}{\partial x}  (x) \cdot \dfrac{\partial}{\partial x} (y) - \dfrac{\partial}{\partial x} (y^2) + \dfrac{\partial}{\partial x} (2) \cdot \dfrac{\partial}{\partial x} (x)  + \dfrac{\partial}{\partial x} (y),$$
$$y - 2x + 2.$$

With respect to y:

$$\dfrac{\partial}{\partial y} (-x^2 + xy - y^2 + 2x  + y),$$
$$= \dfrac{\partial}{\partial y} (-x^2) + \dfrac{\partial}{\partial y}  (x) \cdot \dfrac{\partial}{\partial y} (y) - \dfrac{\partial}{\partial y} (y^2) + \dfrac{\partial}{\partial y} (2) \cdot \dfrac{\partial}{\partial y} (x)  + \dfrac{\partial}{\partial y} (y),$$
$$= x - 2y + 1.$$

Therefore, the gradient is:
$$\nabla f(xy) = \begin{bmatrix} y - 2x + 2 \\ x - 2y + 1 \end{bmatrix}.$$

### (b)
**Work:**

start by solving for x:

$$-2x + y + 2 = 0,$$
$$y = 2x - 2.$$
Plug into the bottom element:
$$x - 2y + 1 = 0,$$
$$x - 2(2x - 2) + 1 = 0,$$
$$x - 4x - 3 = 0,$$
$$3x = 5,$$
$$x = \dfrac {5}{3}.$$
Back to the top element:
$$y - 2x + 2 = 0,$$
$$-y = 2x + 2,$$
$$-y = -2(\frac{5}{3}) + 2,$$
$$y = \dfrac {4}{3}$$


### (c)
**Work:**\newline
The gradient is:
$$\nabla f(xy) = \begin{bmatrix} y - 2x + 2 \\ x - 2y + 1 \end{bmatrix}.$$

Hessian: 
$$H (f(x,y)) = \begin{bmatrix} \dfrac{\partial}{\partial x}(y - 2x + 2) & \dfrac{\partial }{\partial y} (y - 2x + 2) \\ \dfrac{\partial }{\partial x} (x - 2y + 1) & \dfrac{\partial}{\partial y}(x - 2y + 1) \end{bmatrix},$$
$$ = \begin{bmatrix} -2 & 1 \\ 1 & -2 \end{bmatrix}.$$

### (d)
**Work:**
$$f_{xx} (\dfrac {5} {3}, \dfrac {4}{3}) \cdot f_{yy} (\dfrac {5} {3}, \dfrac {4}{3}) - f_{xy} (\dfrac {5} {3}, \dfrac {4}{3})^2 > 0,$$
$$-2(-2) - 1^2 > 0,$$
$$3.$$
We therefore know the matrix is either a positive-definite or negative-definite matrix.

$$f_{xx} (\dfrac {5} {3}, \dfrac {4}{3}) + f_{yy} (\dfrac {5} {3}, \dfrac {4}{3}),$$
$$= -2 + (-2),$$
$$= -4.$$
At $(\dfrac {5} {3}, \dfrac {4}{3})$, the matrix is negative-definite, and at a local maximum. 

## 7.7
**Work:** \newline
First, find gradient of the function (partial derivative with respect to x, then y). With respect to x:
$$\dfrac {\partial}{\partial x} (150x^{1/3} y^{2/3}),$$
$$= 150 y^{2/3}\dfrac {\partial}{\partial x} (x^{1/3}),$$
$$= 150 y^{2/3}\dfrac {1}{3} (x^{-2/3}),$$
$$= 150 y^{2/3}\dfrac {1}{3} \dfrac{1}{x^{2/3}},$$
$$= \dfrac {150y^{2/3}}{3x^{2/3}},$$
$$= \dfrac {50y^{2/3}}{x^{2/3}}.$$
Partial respect to y is identical steps except isolating the y instead:
$$\dfrac{100x^{1/3}}{y^{1/3}}.$$
Which makes the gradient:
$$\nabla f(x,y) = \begin{bmatrix} \frac {50y^{2/3}}{x^{2/3}} \\ \frac{100x^{1/3}}{y^{1/3}} \end{bmatrix}.$$
Step 2 is to find the gradient of the constraint fuction 300x + 500y = 100,000. Partial with respect to x:
$$\dfrac{\partial}{\partial x} (300x + 500y),$$
treating y as constant, derivative of constant (y) = 0:
$$=\dfrac{\partial}{\partial x} (300x),$$
$$=300 \dfrac{\partial}{\partial x} (x),$$
$$=300.$$
Partial with respect to y = 500. Therefore,
$$\nabla g(x,y) = \begin{bmatrix} 300 \\ 500 \end{bmatrix}.$$
Step 3, set each $\nabla f(x,y) = \lambda \nabla g(x,y)$, and Step 4 is to add the constraint to the system:
$$\frac {50y^{2/3}}{x^{2/3}} = 300 \lambda$$
$$\frac{100x^{1/3}}{y^{1/3}} = 500 \lambda$$ 
$$300x + 500y = 100,000$$
If you can make the first two equal to lambda, they will be equal to each other:
$$\frac {50y^{2/3}}{x^{2/3}} = 300 \lambda,$$
$$\frac {50y^{2/3}}{300x^{2/3}} = \lambda,$$
$$\frac {y^{2/3}}{6x^{2/3}} = \lambda.$$
$$\frac{100x^{1/3}}{y^{1/3}} = 500 \lambda,$$ 
$$\frac{100x^{1/3}}{500y^{1/3}} = \lambda,$$ 
$$\frac{x^{1/3}}{5y^{1/3}} = \lambda.$$ 
Now it's a lot easier to solve for x and y:
$$\frac {y^{2/3}}{6x^{2/3}} =\frac{x^{1/3}}{5y^{1/3}},$$ 
$$y^{2/3} =\frac{x^{1/3}}{5y^{1/3}} 6x^{2/3},$$ 
$$5y^{1/3} (y^{2/3}) =x^{1/3} (6x^{2/3}),$$ 
$$5y=6x,$$ 
$$\dfrac{5}{6}y=x.$$ 
$$300(\dfrac{5}{6}y) + 500y = 100,000,$$
$$\dfrac{1500y}{6} + 500y = 100,000,$$
$$250y + 500y = 100,000,$$
$$750y = 100,000,$$
$$y = 133.33.$$
Solving for x:
$$\dfrac{5}{6}y = x,$$
$$\dfrac{5}{6}133.33 = x,$$
$$111.11 = x.$$

Step 6 is to compare f at the critical points. But plugging in 0s does not make equations equal to zero because we can't divide by zero (e.g. plugging zero into $y^{1/3}$ would equal zero). But you can try different points to see how the outcomes of those equations compare to the current one. The current one:
```{r}
150 * 111.11^(1/3) * 133.33^(2/3)
```
Plug in a higher x value (x = 120) and solve for y:
```{python}
from sympy import solveset, S
from sympy.abc import y
from sympy import Symbol, Eq
print(solveset(Eq(300 * 120 + 500*y, 100000),y))
```
Plug in (120, 128) to production function:
```{r}
150 * 120^(1/3) * 128^(2/3)
```
We lose money. What if you decease x? Make x = 100 and solve for y:
```{python}
from sympy import solveset, S
from sympy.abc import y
from sympy import Symbol, Eq
print(solveset(Eq(300 * 100 + 500*y, 100000),y))
```
```{r}
150 * 100^(1/3) * 140^(2/3)
```
Because decreasing x and increasing y (and vice versa) decreases the output of the function, we can be confident that (111.11, 133.33) maximizes the function.

## 7.10

### (a)
**Work:** \newline
We already have everything we need: \newline
$s(X) = \sqrt{4} = 2$ \
$s(Y) = \sqrt {9} = 3$ \
$$Corr(X,Y) = \dfrac {Cov(X,Y)}{s(X) \cdot s(Y)},$$
$$= \dfrac {5}{6},$$
$$= 8.33.$$

### (b)
**Work:** \newline
First remove constant since V(X + C) = V(X):
$$V(A) = V(3X - 3Y),$$
$$V(A) = V(3X - 3Y),$$
Weighted difference, so use the weighted sum of two variances formula (which is the same as for when you need a difference also):
$$V(aX - bY) = a^2V(X) + b^2V(Y) + 2ab\text{Cov}(X,Y),$$
$$ = 3^2V(X) + (-3)^2V(Y) + 2(3)(-3)\text{Cov}(X,Y),$$
Then plug in the information the question has given:
$$ = 3^2(4) + (-3)^2(9) + 2(3)(-3)(5),$$
```{r}
3^2 * 4 + (-3)^2 * 9 + 2 * 3 * (-3) * 5
```

### (c)
**Work:** \newline
First remove constant since V(X + C) = V(X):
$$V(B) = V(-2X + Y),$$
Weighted difference, so use the weighted sum of two variances formula (which is the same as for when you need a difference also):
$$V(aX - bY) = a^2V(X) + b^2V(Y) + 2ab\text{Cov}(X,Y),$$
$$ = -2^2V(X) + (1)^2V(Y) + 2(-1)(1)\text{Cov}(X,Y),$$
Then plug in the information the question has given:
$$ = (-2)^2(4) + 1^2(9) + 2(-1)(1)(5).$$
```{r}
(-2)^2 * 4 + 1^2 * 9 + 2 * (-1) * (1) * 5
```

### (d)
**Work:** \newline
When both variables are replaced by linear functions of other variables, you use:
$$\text{Cov}(aX + bY, cW + dZ) = ac\text{Cov}(X, W) + ad\text{Cov}(X, Z) + bc\text{Cov}(Y, W) + bd\text{Cov}(Y, Z)$$
$$= 3(-2)\text{Cov}(X, W) + 3(1)\text{Cov}(X, Z) + (-3)(-2)\text{Cov}(Y, W) + (-3)(1)\text{Cov}(Y, Z),$$
$$= 3(-2)(4) + 3(1)(5) + (-3)(-2)(5) + (-3)(1)(9),$$
```{r}
3*(-2)*(4) + 3*(1)*(5) + (-3)*(-2)*(5) + (-3)*(1)*(9)
```

### (e)
**Work:** \newline
$$\dfrac{-6}{\sqrt{27}\sqrt{15}}.$$
```{r}
-6/(sqrt(27) * sqrt(15))
```
