---
title: "CatData HW5"
author: "Matthew Vanaman"
date: '`r format(Sys.Date(), "%m-%d-%Y")`'
monofont: Courier New
output:
  pdf_document: 
    latex_engine: xelatex
  word_document: default
header-includes: \usepackage{tikz,xcolor,listings}
---
`r knitr::opts_knit$set(root.dir='..')`

```{r, include=FALSE}
library(ProjectTemplate); load.project()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
options(scipen = 999)
knitr::opts_chunk$set(engine.path = "/anaconda3/bin/python")
np <- reticulate::import("numpy")
sympy <- reticulate::import_from_path("sympy", path = "/anaconda3/lib/python3.7/site-packages")
```

3,20, 3,21, 3.22

# 3.19

## (a)

```{r, echo=FALSE}
trainData <- as.data.frame(matrix(c(2003, 2002, 2001, 2000, 1999, 1998, 1997, 1996, 1995, 
                      1994, 1993, 1992, 1991, 1990, 1989, 1988, 1987, 1986, 
                      1985, 1984, 1983, 1982, 1981, 1980, 1979, 1978, 1977, 
                      1976, 1975,
                      
                      518, 516, 508, 503, 505, 487, 463, 437, 423, 415, 
                      425, 430, 439, 431, 436, 443, 397, 414, 418, 389, 
                      401, 372, 417, 430, 426, 430, 425, 426, 436,
                      
                      0, 1, 0, 1, 1, 0, 1, 2, 1, 2, 
                      0, 1, 2, 1, 4, 2, 1, 2, 0, 5, 
                      2, 2, 2, 2, 3, 2, 1, 2, 5,
                      
                      3, 3, 4, 3, 2, 4, 1, 2, 2, 4, 
                      4, 4, 6, 2, 4, 4, 6, 13, 5, 3, 
                      7, 3, 2, 2, 3, 4, 8, 12, 2), 
                    
                      nrow = 29, ncol = 4))
colnames(trainData) <- c("Year", "TrainKm", 
                         "TrainCollisions", "TrainRoadCollisions")
```

```{r, echo=FALSE}
mod_intercept_only <- glm(TrainCollisions ~ 1, offset = log(TrainKm), data=trainData, family = poisson)
mod_with_time <- glm(TrainCollisions ~ Year, offset = log(TrainKm), data=trainData, family = poisson)
```
Confirm the sizes of the deviances:
```{r, echo=FALSE}
mod_intercept_only$formula
mod_intercept_only$deviance
mod_with_time$formula
mod_with_time$deviance
```
Get the difference in deviances and residual df between the models:
```{r}
mod_intercept_only$deviance - mod_with_time$deviance
mod_intercept_only$df.residual - mod_with_time$df.residual
```
The deviance approximates a chi-squared distribution when samples are large. Get the p-value for $\Delta \chi^2$:
```{r}
pchisq(11.60185, df=1, lower.tail = FALSE)
```


## (b)
Get the likelihood $\chi^2$, or $z^2$, and p-value with 1 degree of freedom:
```{r}
(-0.0337/ 0.0130)^2
pchisq(6.720059, df=1, lower.tail = FALSE)
```

## (c)

The confidence intervals from the negative binomial model are around an additive, linear effect (because negative binomials are a special mixture of Poisson distributions). To get confidence intervals around a multiplicative effect, you exponentiate the intervals:
```{r}
c(exp(-0.060), exp(-0.008))
```


# 3.20

## (a)

```{r, echo=FALSE}
deadData <- as.data.frame(matrix(c("35-44", "45-54", "55-64", "65-74", "75-84",
                     18793, 10673, 5710, 2585, 1462,
                     52407, 43248, 28612, 12663, 5317,
                     2, 12, 28, 28, 31,
                     32, 104, 206, 186, 102), nrow = 5, ncol = 5))
colnames(deadData) <- c("Age", "PY_nonSmokers", "PY_Smokers", "CD_nonSmokers", "CD_Smokers")
deathrates <- as.data.frame(matrix(c("35-44", "45-54", "55-64", "65-74", "75-84",
                       2/(18793/1000), 12/(10673/1000), 28/(5710/1000), 28/(2585/1000), 
                       31/(1462/1000),
                       
                       32/(52407/1000), 104/(43248/1000), 206/(28612/1000),
                       186/(12663/1000), 102/(5317/1000),
                       
                       (2/(18793/1000))/(32/(52407/1000)), (12/(10673/1000)/(104/(43248/1000))),
                       (28/(5710/1000)/(206/(28612/1000))), (28/(2585/1000)/(186/(12663/1000))), 
                       (31/(1462/1000)/(102/(5317/1000)))), 
                       ncol = 4, nrow = 5))
colnames(deathrates) <- c("Age", "DeathRate_nonSmokers", "DeathRate_Smokers", "Ratio")
deathrates
```

Deathrates for smokers and non-smokers alike increase with age, as one would expect. Of interest is the fact that the ratio of death rats also increases, such that as people age, the death rates between those who smoke and those who don't approaches a 1:1 ratio (which would mean no effect of smoking on death rate beyond age) before exceeding a 1:1 ratio after age 75. This implies curvilinear relationship between smoking and age. You stand greater risk of death up until you are 75, but if you make it to 75 you may benefit from being smoker.



```{r}
class(deathrates$DeathRate_nonSmokers)
```

