---
title: "CatData HW6"
author: "Matthew Vanaman"
date: '`r format(Sys.Date(), "%m-%d-%Y")`'
monofont: Courier New
output:
  pdf_document: 
    latex_engine: xelatex
  word_document: default
header-includes: \usepackage{tikz,xcolor,listings}
---
`r knitr::opts_knit$set(root.dir='..')`

```{r, include=FALSE, cache=TRUE}
library(ProjectTemplate); load.project()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
options(scipen = 999)
knitr::opts_chunk$set(engine.path = "/anaconda3/bin/python")
np <- reticulate::import("numpy")
sympy <- reticulate::import_from_path("sympy", path = "/anaconda3/lib/python3.7/site-packages")
```

A4.5, A4.15 (3 points)

A4.31 (5 points)

A4.16-A4.17, A5.4 (10 points)




# 4.3

**\large (a):** In the linear probability model, the coefficient is treated the same way as it is in linear regression. That is, it represents a slope such that for every unit increase x (decade, in this case), there is a corresponding unit change in y (the probability of pitching a complete game, in this case, the change of which = -0.0694).

**\large (b):** 
$$\hat \pi = 0.7578 - 0.0694(12),$$
$$= `r 0.7578 - 0.0694*(12)`.$$
Percentages can't be negative! So not possible.

**\large (c):** 
$$\hat \pi = \dfrac {e^{\alpha + \beta x}} {1 + e^{\alpha + \beta x}}.$$
$$\hat \pi = \dfrac {e^{1.148 - 0.315 (12)}} {1 + e^{1.148 - 0.315 (12)}},$$
$$= `r exp(1)^(1.148 - 0.315 * 12) / (1 + exp(1)^(1.148 - 0.315 * 12))`.$$
This is much more plausible. 


# 4.5






