---
title: "CatData HW6"
author: "Matthew Vanaman"
date: '`r format(Sys.Date(), "%m-%d-%Y")`'
monofont: Courier New
output:
  pdf_document: 
    latex_engine: xelatex
  word_document: default
header-includes: \usepackage{tikz,xcolor,listings}
---
`r knitr::opts_knit$set(root.dir='..')`

```{r, include=FALSE, cache=TRUE}
library(ProjectTemplate); load.project()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
options(scipen = 999)
knitr::opts_chunk$set(engine.path = "/anaconda3/bin/python")
np <- reticulate::import("numpy")
sympy <- reticulate::import_from_path("sympy", path = "/anaconda3/lib/python3.7/site-packages")
```

A4.3, A4.5, A4.15 (3 points)

A4.31 (5 points)

A4.16-A4.17, A5.4 (10 points)




# 4.3

### (a): 
*Show how software obtained $\hat \pi$ = 0.068 when x = 8.*

Software calculated predicted probabilities at each level of the independent variable, including x = 8. In logistic regression, the following formula yields the estimate:
$$\hat \pi = \dfrac {e^{\alpha + \beta x}} {1 + e^{\alpha + \beta x}}.$$
Table 4.9 contains the coeficient and y-intercept. Plugging those in, you get:
$$\hat \pi = \dfrac {e^{-3.7771 + 0.1449 (8)}} {1 + e^{-3.7771 + 0.1449 (8)}},$$
$$= `r exp(1)^(-3.7771 + 0.1449 * 8) / (1 + exp(1)^(-3.7771 + 0.1449  * 8))`.$$

### (b)
*Show that $\hat \pi$ = 0.50 when x = 26.0.*

Same deal as with part a:
$$\hat \pi = \dfrac {e^{\alpha + \beta x}} {1 + e^{\alpha + \beta x}}.$$
$$\hat \pi = \dfrac {e^{-3.7771 + 0.1449 (26.0)}} {1 + e^{-3.7771 + 0.1449 (26.0)}},$$
$$= `r exp(1)^(-3.7771 + 0.1449 * 26.0) / (1 + exp(1)^(-3.7771 + 0.1449  * 26.0))`.$$

### (c)
*Show that the rate of change in $\hat \pi$ is 0.009 when x = 8 and is 0.036 when x = 26.*

Aha! To get the instantaneous rate of change formula, you take the first derivative with respect to x. So you have:
$$\hat \pi = \dfrac {e^{\alpha + \beta x}} {1 + e^{\alpha + \beta x}}.$$
Then the first dervivative comes out to be:
$$f'(x) = \dfrac {\beta \ e^{\alpha + \beta x}} {(1 + e^{\alpha + \beta x})^2},$$
$$\dfrac {0.1449 (e^{-3.7771 + 0.1449(26)})} {(1 + e^{-3.7771 + 0.1449(26)})^2},$$
$$= `r (0.1449 * exp(1)^(-3.7771 + 0.1449*26)) / (1 + exp(1)^(-3.7771 + 0.1449*26))^2`.$$
This formula is the probability density function of the logistic distribution. Removing the exponent in the denominator yields the logistic comulative distribution function. \newline

Repeat for the other one:
$$\dfrac {0.1449 (e^{-3.7771 + 0.1449(8)})} {(1 + e^{-3.7771 + 0.1449(8)})^2},$$
$$= `r (0.1449 * exp(1)^(-3.7771 + 0.1449*8)) / (1 + exp(1)^(-3.7771 + 0.1449*8))^2`.$$

### (d)






