---
title: "Math HW5"
author: "Matthew Vanaman"
date: '`r format(Sys.Date(), "%m-%d-%Y")`'
monofont: Courier New
output:
  pdf_document: 
    latex_engine: xelatex
  word_document: default
header-includes: \usepackage{tikz,xcolor,listings}
---
`r knitr::opts_knit$set(root.dir='..')`

```{r, include=FALSE}
library(ProjectTemplate); load.project()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
options(scipen = 999)
knitr::opts_chunk$set(engine.path = "/anaconda3/bin/python")
np <- reticulate::import("numpy")
sympy <- reticulate::import_from_path("sympy", path = "/anaconda3/lib/python3.7/site-packages")
```

4.8, 4.9, 5.1, 5.2, 5.3

# 4.8
## (a)



### Work 

#### First Derivative:
The only layer here is in the exponent of $e$; everything else is constant.
$$f (x) = \dfrac {1} {\sqrt{2 \pi}} e^{-\frac {x^2}{2}}, $$
$$A = -\frac {x^2}{2}, \ \ f(x) = \dfrac {1} {\sqrt{2 \pi}} e^A $$
$$A' = \dfrac {d}{dx} \Big( -\dfrac{x^2}{2} \Big) \\ = - \dfrac {1}{2} \Big(2x^{2-1}) \\ = - \dfrac {2x^{2-1}}{2} \\ = -x. $$
$$ \dfrac {1} {\sqrt{2 \pi}} e^A \Big(-x \Big) = \dfrac {-x} {\sqrt{2 \pi}} e^{-\frac {x^2}{2}}. $$
#### Second Derivative

Because the first derivative was multiplicative, we need the product to get the second derivative. \newline
Product rule: $f''(x) = g'(x) f(x) + g(x) f'(x).$ \newline
$f(x) =  \dfrac {1} {\sqrt{2 \pi}} e^{-\frac {x^2}{2}}.$ \newline
$f'(x) = \dfrac {-x} {\sqrt{2 \pi}} e^{-\frac {x^2}{2}}.$ \newline
$g(x) = -x. $
$g'(x) = \frac {d}{dx}(-x) = -1 .$

$$f''(x) = \bigg( \dfrac {x^2} {\sqrt{2 \pi}} e^{-\frac {x^2}{2}}\bigg) - \bigg( \dfrac {1} {\sqrt{2 \pi}} e^{-\frac {x^2}{2}} \bigg) \\ = \bigg( \dfrac {x^2 - 1} {\sqrt{2 \pi}} e^{-\frac {x^2}{2}}\bigg). $$

#### Plug in 0 for each version

$$f(0) = \dfrac {1} {\sqrt{2 \pi}} e^{-\frac {0^2}{2}} = \dfrac {1}{\sqrt{2 \pi}}. $$
$$f'(0) = \dfrac {0^2} {\sqrt{2 \pi}} e^{-\frac {0^2}{2}} = 0. $$
$$f''(0) = \dfrac {0^2 - 1} {\sqrt{2 \pi}} e^{-\frac {0^2}{2}} = \dfrac {-1} {\sqrt{2 \pi}}. $$
Taylor approximation - just plug in what we found:
$$f(x) \approx \dfrac {1}{\sqrt{2 \pi}} + 0x + \dfrac {-1 / \sqrt{2 \pi}} {2}x^2.$$
Which if we clean up and rearrange to the more recognizable quadratic form, looks like:
$$f''(x) \approx \dfrac {-1}{\sqrt{8 \pi}}x^2 + \dfrac {1} {\sqrt{2 \pi}}. $$

## (b)

```{r, echo=FALSE}

```

